A Phenomenological Model for the Intracluster Medium that matches X-ray and Sunyaev-Zel'dovich observations

### 2013-10-11 paper1.tex
dereddened text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf Cosmological hydrodynamical simulations of galaxy clusters are still
    challenged to produce a model for the intracluster medium that matches all
    aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
    such comparisons with future simulations and to enable realistic cluster
    population studies for modeling e.g., non-thermal emission processes, we
    construct a phenomenological model for the intracluster medium that is based
    on a representative sample of \qc{observed} X-ray clusters. We create a mock galaxy
    cluster catalog \qc{based on the large collisionless $N$-body simulation MultiDark, by 
    assigning our gas density model to each dark matter cluster halo.}
    %by assigning our gas density model to each cluster halo of
    %the large collisionless $N$-body simulation MultiDark. 
    Our clusters \qc{are classified as cool-core and non cool-core} according to a
    dynamical disturbance parameter. We demonstrate that our gas model matches
    the various observed Sunyaev-Zel'dovich and X-ray scaling relations as well
    as the X-ray luminosity function, thus enabling \qc{to build} a reliable mock catalog for
    present surveys and forecasts for future \qc{experiments}. In a companion paper,
    we apply our catalogs to calculate non-thermal radio and gamma-ray emission
    of galaxy clusters. We make our cosmologically complete multi-frequency mock
    catalogs for the (non-)thermal cluster emission at different redshifts
    publicly and freely available online through the MultiDark database.}
\end{abstract}

@@

  The central cooling time of the X-ray emitting gas in galaxy clusters and
  groups is bimodally distributed. Approximately half of all systems have
  radiative cooling times of less than 1~Gyr and establish a population of cool
  core clusters (CCCs)\qc{, while} others have cooling times that can be longer than
  the age of the Universe, forming the non-cool core cluster (NCCC) population
  \citep{2009ApJS..182...12C,2010A&A...513A..37H}. Several physical processes
  have been proposed to be responsible for balancing radiative cooling of the
  low-entropy gas at the centers of CCCs. In particular feedback by active
  galactic nuclei (AGN) has come into the focus of recent research
  \citep{2007ARA&A..45..117M, 2012NJPh...14e5023M}\qc{, owing} to the self-regulated
  nature of the proposed feedback and observational correlations between radio
  activity and central entropy (which is a proxy for a small cooling time). The
  underlying physical processes responsible for the heating include dissipation
  of mechanical heating by outflows, lobes, or sound waves from the central AGN
  \citep[e.g.,][]{2001ApJ...554..261C, 2002Natur.418..301B, 2002ApJ...581..223R,
    2012MNRAS.424..190G} or cosmic-ray heating: streaming cosmic rays excite
  Alfv\'en waves on which they resonantly scatter \citep{1969ApJ...156..445K}.
  Damping of these waves transfers cosmic ray energy and momentum to the cooling
  plasma \citep{1991ApJ...377..392L, 2008MNRAS.384..251G, 2011A&A...527A..99E,
    2013MNRAS.434.2209W}, heating it at a rate that balances radiative cooling
  on average at each radius while explaining the observed temperature floor in
  CCCs by thermal stability of the heating mechanism
  \citep{2013arXiv1303.5443P}.

  Modeling the formation and evolution of clusters with cosmological
  hydrodynamical simulations has been progressively refined over the last years
  to also include (sub-resolution models for) AGN feedback
  \citep[e.g.,][]{2007MNRAS.380..877S, 2008MNRAS.387.1403S, 2008ApJ...687L..53P,
    2012MNRAS.420.2662D, 2012MNRAS.424..190G, 2013MNRAS.428.2366V}. While the
  \qc{comparison between} data and simulations with AGN feedback improved for integrated
  thermal properties of large cluster samples
  \citep{2012ApJ...758...74B,2012ApJ...758...75B}, there are still notable
  differences for differential quantities such as the pressure profile
  \citep{2013A&A...550A.131P} or the profile of gas mass fractions
  \citep{2012arXiv1209.4082B}. These discrepancies (that have been more dramatic
  in the past without AGN feedback) motivated the development of
  phenomenological models of the ICM \citep[e.g.,][]{2005ApJ...634..964O,
    2012MNRAS.422..686C}. Here, we construct a simple and purely
  phenomenological model that matches all available X-ray and Sunyaev-Zel'dovich
  (SZ) data with the goal to facilitate \qc{comparison with} future hydrodynamical
  simulations, to allow the modeling of non-thermal cluster emission over the
  entire electromagnetic \qc{wave-band}, and finally to enable the construction of
  cluster mock catalogs of the thermal and non-thermal cluster emission (in the
  radio, hard X-ray, and gamma-ray band) of current and future surveys. Such a
  model for the ICM needs to be applied to a sample of cluster halos, which
  can either be obtained by means of analytical expressions for the cluster mass
  function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological
  simulations. In this work, we make use of the recent \qc{large-volume} $N$-body
  simulation MultiDark, which employs a \qc{flat $\Lambda$ Cold Dark Matter} cosmology
  \citep{2011arXiv1104.5130P}.

  Our \emph{phenomenological} approach uses gas density profiles of a
  representative sample of \qc{observed} X-ray clusters, which are complemented by 
  \qc{the observed} cluster mass-dependent gas mass fractions. The resulting cluster mock sample
  are then compared to data including the X-ray luminosity function (XLF), the
  luminosity-mass relation, $L_{\rmn{X}}- M$, and the $Y_{\rmn{X}}-M$ relation,
  where $Y_{\rmn{X}}=M_\rmn{gas}k_{\rmn{B}}T$ with an X-ray-derived gas mass
  $M_{\rmn{gas}}$ and temperature $T$ \citep{2006ApJ...650..128K}. We
  additionally compare our $Y_{\rmn{SZ}}-M$ relation to recent SZ data, where
  $Y_\rmn{SZ}$ denotes the integrated Compton-$y$ parameter.  In a companion
  paper (\citealp{paper2}; hereafter Paper~II), we apply our ICM model to
  predict the non-thermal \qc{radio and gamma-ray} emission of a cosmological complete 
  sample of galaxy clusters, enabling valuable insight into the statistics of radio halos.
  One of our final data products is a complete cosmological cluster mock catalog,
  at different redshifts, containing information of \qc{the} DM and ICM properties for
  each object, together with its (non-)thermal emission at different
  frequencies. \qc{All the products can be found on-line at the MultiDark 
  database (\textit{www.multidark.org}).}

  In Section~\ref{sec:2}, we introduce the MultiDark simulation along with our
  \qc{selected} cluster halo sample. We assign to each of our clusters the
  phenomenologically constructed gas density profile in Section~\ref{sec:3}. In
  Sections~\ref{sec:4} and \ref{sec:5}, we show that this approach can
  successfully reproduce the observed X-ray and SZ cluster data. We present the
  resulting mock cluster catalogs in Section~\ref{sec:6}, and eventually
  summarize in Section~\ref{sec:7}.  In this work, the cluster mass $M_{\Delta}$
  and radius $R_{\Delta}$ are defined with respect to a density that is
  $\Delta=200$ or 500 times the \emph{critical} density of the Universe. We
  adopt density parameters of $\Omega_{\rmn{m}}=0.27$, $\Omega_{\Lambda}=0.73$ and
  today's Hubble constant of $H_0 = 100\, h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where
  $h_{70} = 0.7$.}

@@

{\bf In Paper~II, we present predictions for the LOFAR radio observatory
which \qc{expects} to detect diffuse radio emission in clusters up to redshift $z \approx 1$ (e.g.,
\citealp{2012JApA..tmp...34R}).} Thus, we make use of different simulation snapshots up to $z=1$. 
{\bf The extrapolation of our model beyond this redshift is rather uncertain as it is based
on observations at low(er) redshift.} In Table~\ref{tab:z}, we show the total cluster number in 
our final cluster sample at different redshifts.

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/lx_m.eps}
\includegraphics[width=0.48\textwidth]{figures/yx_m.eps}
\caption{{\bf X-ray \qc{cluster} scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result by \protect\cite{2010MNRAS.406.1773M}. 
  The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin. The bottom panels show the relative difference to the observational
    scaling relations. \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}.}} 
\label{fig:X_LM}
\end{figure*}

@@

{\bf So far, we used a well-observed representative sample of X-ray clusters
(REXCESS) that was supplemented \qc{with} X-ray data on gas mass fractions to construct
our phenomenological gas model.  Here, we will test how our cluster mock data
compare to various X-ray inferred observables, which are obtained from different
cluster samples and from (partially) different X-ray observatories. Those
include the observed $L_{\rmn{bol}} - M_{500}$ relation, the $Y_X - M_{500}$
relation, and the XLF.}

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m.eps}
\includegraphics[width=0.48\textwidth]{figures/sz_m_ACT.eps}
\caption{{\bf Comparison of our \qc{MultiDark} SZ scaling relations (grey triangles) with
    data. The black crosses represent
    the $Y_{\rmn{SZ}}$ median values of our sample for a given mass bin
    (indicated by horizontal error bars), and the vertical error bars represent
    the standard deviation within a bin. 
    The bottom panels show the relative difference to the observational
    scaling relations.
    \emph{Left.}  We compare our $Y_{\rmn{SZ}}-M$ relation at
    $z=0.18$ with the {\em Planck} data \citep[blue,][]{2013arXiv1303.5080P},
    which have a median redshift of about $0.18$. \emph{Right.} We compare our
    $Y_{\rmn{SZ}}-M$ relation at $z=0.53$ with the ACT sample
    \protect\citep[blue,][]{2013JCAP...07..008H} assuming a self-similar
    redshift evolution.}}
\label{fig:SZ_M}
\end{figure*}

In the left panel of Fig.~\ref{fig:SZ_M}, we compare the $Y_{\rmn{SZ}}-M_{500}$
relation of our model (following \citealp{2012ApJ...758...74B}, equation~3) with
the observed scaling relation by the \cite{2013arXiv1303.5080P}. {\bf We use the
  \emph{Planck} COSMO sample, which is Malmquist-bias corrected and has a median
  redshift of $0.18$, and compare this to our $z=0.18$ relation}. Our model
reproduces the normalization of the scaling relation remarkably well, except for
the high-mass end where our simulations have a weaker constraining power due to
the smaller box size in comparison to the survey volume of {\em Planck}.  {\bf
  However, the slope of \qc{the MultiDark} $Y_{\rmn{SZ}}-M_{500}$ relation ($1.59\pm0.01$) is
  shallower in comparison to the self-similar slope ($5/3$) as well as the
  observed slope of the \emph{Planck} COSMO sample. We can analytically
  understand the cluster-mass scaling of our model by considering its
  mass-dependent quantities, namely $Y_{\rmn{SZ}}\propto M_{500} f_{\rmn{gas}}
  T \propto M_{500}^{1.595}$. In particular, we can trace back the
  shallower slope to the adopted temperature scaling of our model
  ($T\approx T_{\rmn{ci}}\propto M_{500}^{0.46}$), which includes the cluster core region
  and yields a shallower mass scaling in comparison to the self-similar
  expectation ($T\propto M_{500}^{2/3}$). The temperature scaling is only
  partially countered by the scaling of the gas mass fraction with cluster mass,
  $f_{\rmn{gas}}\propto M_{500}^{0.135}$. To improve upon this, we would have to
  account for a slightly steeper temperature-mass scaling \qc{relation} (e.g., using a
  centrally excised temperature scaling, see \citealt{2010MNRAS.406.1773M}) in
  the outer cluster parts that are of relevance for the SZ flux, as explained in
  Sect.~\ref{sec:Xscaling}. }  We find a scatter of $\sigma_{yx} \approx 0.11$
which compares well with the \emph{Planck} result of $\sigma_{yx} \approx
0.08$. In Table~\ref{tab:YSZfits}, we report our SZ scaling relations for
different redshifts.

@@

{\bf In the right panel of Fig.~\ref{fig:SZ_M}, we compare our scaling relation
  to the recent Atacama Cosmology Telescope (ACT) sample
  \citep{2013JCAP...07..008H}, which has a median redshift of $z=0.53$.  Again,
  the slope of our modeled scaling relation is \qc{slightly} shallower in comparison
  to the observed slope. Most strikingly, there appears to be a 30 per cent
  offset of the scaling relations, which could be due to a number of reasons. On
  the data side, these could include possible selection effects of the different
  cluster samples or possible biases associated with density clumping and/or
  non-thermal pressure contributions inherited from the X-ray informed
  post-processing that was employed by the ACT Collaboration to construct their
  $Y_{500}$ estimates. This could also indicate that our model is too
  simplistic to capture the rich cluster physics and potentially may even \qc{point to}
  deviations from the assumed self-similar redshift evolution (since the ACT
  sample probes on average higher redshift and less massive clusters in
  comparison to the {\em Planck} sample).  However, note that our scaling
  relation at $z=0.53$ ($A=-27.67\pm0.12$, $B=1.60\pm0.01$) is within the
  uncertainties compatible with that of the ACT sample ($A=-28.59\pm2.26$,
  $B=1.65\pm0.15$).  Thus, we conclude that overall there is reasonable
  agreement between our phenomenological model and the SZ data of the
  \emph{Planck} and ACT collaborations.}

@@

{\bf As a final product of this work, we construct cosmologically complete mock
  catalogs of galaxy clusters at different redshifts (see Table~\ref{tab:z}) for
  $M_{200}\geq1\times10^{14}$~$h^{-1}\rmn{M}_{\odot}
  \approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$.  We make these catalogs
  publicly available through the MultiDark database \qc{(\textit{www.multidark.org})} where we
  will also post possible updates.

  Our catalogs contain all the information regarding the DM properties of each
  \qc{cluster} as given by the corresponding original MultiDark BDM halo
  catalogs. Additionally, we include the following information:

@@

In Paper~II, we present a model \qc{that allow us to compute the possible radio 
and gamma-ray emission for each cluster in our mock catalog},
%to calculate the possible radio and gamma-ray emission of clusters 
as result of the cosmic-ray (CR) proton-proton interactions
with the ICM. We refer to that paper for all the details. However, we want to
point out that our mock catalogs also include:

@@

We note that our mock catalogs are not appropriate for all purposes because the
density \qc{profiles} that we adopt in our model represent the average cluster
population of cool-core and non cool-core systems (while we account for scatter
in the gas mass fractions, i.e., the normalization of the profiles). As such,
\qc{our mocks} are very useful as a baseline for comparisons to new hydrodynamical
simulations and to new observational samples\qc{, as} well as for non-thermal modeling
of clusters. On the contrary, \qc{in the current form they are not} appropriate to perform detailed
cosmology studies. Careful modeling of the response function of the considered
instrument, in combination with light-cone mock catalogs, would be required in
this case. Nevertheless, for \qc{volume-limited cluster samples, our mock catalogs 
will certainly serve as valuable tools. }
%volume-complete surveys, our mock catalogs will certainly serve as valuable tools.  }

@@

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. Note that our model is calibrated
  \qc{with} X-ray observables within $R_{500}$. More work would be needed to
  extrapolate the model to larger cluster radii that are characterized by an
  increased complexity, which manifests itself in a larger kinetic-to-thermal
  pressure support, an increased level of density clumping, as well as cluster
  asphericity and ellipticity. In our model, we assign a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution. Applying the model to our sample
  of cluster halos, we obtain \qc{volume-limited} mock catalogs for the thermal (SZ
  and X-ray) cluster emission. Our mock catalogs match the observed X-ray
  luminosity function as well as the $L_{\rmn{X, bol}}-M_{500}$,
  $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$ relations well.

  However, there are some deviations among the different observational scaling
  relations and our model, which may either point to observational sample
  selection effects that are not accounted for or missing complexity of our
  model. This model was specifically constructed to provide a reliable
  description of the gas density, which necessarily implies a great match of the
  resulting mock X-ray observables to data (scaling relations and luminosity
  functions)\qc{, as well as} to the low-redshift and massive cluster sample probed by
  {\em Planck}. However, we note that our model has a \qc{slightly} shallower slope
  of the $Y_{\rmn{SZ}}-M_{500}$ relation, which implies that the X-ray inferred
  deviations of self-similarity on small scales become weaker on the larger
  scales probed through the SZ effect. Moreover, our model overpredicts the
  normalization of the ACT $Y_{\rmn{SZ}}-M_{500}$ relation (at $z=0.53$) by
  about 30 per cent. While this offset is still consistent within the 1-$\sigma$
  uncertainty intervals of the ACT scaling relation, this could also indicate
  possible selection effects of the different cluster samples or alternatively
  \qc{points to} a deviation from the assumed self-similar redshift evolution.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf Cosmological hydrodynamical simulations of galaxy clusters are still
    challenged to produce a model for the intracluster medium that matches all
    aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
    such comparisons with future simulations and to enable realistic cluster
    population studies for modeling e.g., non-thermal emission processes, we
    construct a phenomenological model for the intracluster medium that is based
    on a representative sample of observed X-ray clusters. We create a mock
    galaxy cluster catalog based on the large collisionless $N$-body simulation
    MultiDark, by assigning our gas density model to each dark matter cluster
    halo.  Our clusters are classified as cool-core and non cool-core according
    to a dynamical disturbance parameter. We demonstrate that our gas model
    matches the various observed Sunyaev-Zel'dovich and X-ray scaling relations
    as well as the X-ray luminosity function, thus enabling to build a reliable
    mock catalog for present surveys and forecasts for future experiments. In a
    companion paper, we apply our catalogs to calculate non-thermal radio and
    gamma-ray emission of galaxy clusters. We make our cosmologically complete
    multi-frequency mock catalogs for the (non-)thermal cluster emission at
    different redshifts publicly and freely available online through the
    MultiDark database.}
\end{abstract}

@@

  The central cooling time of the X-ray emitting gas in galaxy clusters and
  groups is bimodally distributed. Approximately half of all systems have
  radiative cooling times of less than 1~Gyr and establish a population of cool
  core clusters (CCCs), while others have cooling times that can be longer than
  the age of the Universe, forming the non-cool core cluster (NCCC) population
  \citep{2009ApJS..182...12C,2010A&A...513A..37H}. Several physical processes
  have been proposed to be responsible for balancing radiative cooling of the
  low-entropy gas at the centers of CCCs. In particular feedback by active
  galactic nuclei (AGN) has come into the focus of recent research
  \citep{2007ARA&A..45..117M, 2012NJPh...14e5023M}, owing to the self-regulated
  nature of the proposed feedback and observational correlations between radio
  activity and central entropy (which is a proxy for a small cooling time). The
  underlying physical processes responsible for the heating include dissipation
  of mechanical heating by outflows, lobes, or sound waves from the central AGN
  \citep[e.g.,][]{2001ApJ...554..261C, 2002Natur.418..301B, 2002ApJ...581..223R,
    2012MNRAS.424..190G} or cosmic-ray heating: streaming cosmic rays excite
  Alfv\'en waves on which they resonantly scatter \citep{1969ApJ...156..445K}.
  Damping of these waves transfers cosmic ray energy and momentum to the cooling
  plasma \citep{1991ApJ...377..392L, 2008MNRAS.384..251G, 2011A&A...527A..99E,
    2013MNRAS.434.2209W}, heating it at a rate that balances radiative cooling
  on average at each radius while explaining the observed temperature floor in
  CCCs by thermal stability of the heating mechanism
  \citep{2013arXiv1303.5443P}.

  Modeling the formation and evolution of clusters with cosmological
  hydrodynamical simulations has been progressively refined over the last years
  to also include (sub-resolution models for) AGN feedback
  \citep[e.g.,][]{2007MNRAS.380..877S, 2008MNRAS.387.1403S, 2008ApJ...687L..53P,
    2012MNRAS.420.2662D, 2012MNRAS.424..190G, 2013MNRAS.428.2366V}. While the
  comparison between data and simulations with AGN feedback improved for integrated
  thermal properties of large cluster samples
  \citep{2012ApJ...758...74B,2012ApJ...758...75B}, there are still notable
  differences for differential quantities such as the pressure profile
  \citep{2013A&A...550A.131P} or the profile of gas mass fractions
  \citep{2012arXiv1209.4082B}. These discrepancies (that have been more dramatic
  in the past without AGN feedback) motivated the development of
  phenomenological models of the ICM \citep[e.g.,][]{2005ApJ...634..964O,
    2012MNRAS.422..686C}. Here, we construct a simple and purely
  phenomenological model that matches all available X-ray and Sunyaev-Zel'dovich
  (SZ) data with the goal to facilitate comparison with future hydrodynamical
  simulations, to allow the modeling of non-thermal cluster emission over the
  entire electromagnetic wave-band, and finally to enable the construction of
  cluster mock catalogs of the thermal and non-thermal cluster emission (in the
  radio, hard X-ray, and gamma-ray band) of current and future surveys. Such a
  model for the ICM needs to be applied to a sample of cluster halos, which
  can either be obtained by means of analytical expressions for the cluster mass
  function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological
  simulations. In this work, we make use of the recent large-volume $N$-body
  simulation MultiDark, which employs a flat $\Lambda$ Cold Dark Matter cosmology
  \citep{2011arXiv1104.5130P}.

  Our \emph{phenomenological} approach uses gas density profiles of a
  representative sample of observed X-ray clusters, which are complemented by
  the observed cluster mass-dependent gas mass fractions. The resulting cluster
  mock sample are then compared to data including the X-ray luminosity function
  (XLF), the luminosity-mass relation, $L_{\rmn{X}}- M$, and the $Y_{\rmn{X}}-M$
  relation, where $Y_{\rmn{X}}=M_\rmn{gas}k_{\rmn{B}}T$ with an X-ray-derived
  gas mass $M_{\rmn{gas}}$ and temperature $T$ \citep{2006ApJ...650..128K}. We
  additionally compare our $Y_{\rmn{SZ}}-M$ relation to recent SZ data, where
  $Y_\rmn{SZ}$ denotes the integrated Compton-$y$ parameter.  In a companion
  paper (\citealp{paper2}; hereafter Paper~II), we apply our ICM model to
  predict the non-thermal radio and gamma-ray emission of a cosmological
  complete sample of galaxy clusters, enabling valuable insight into the
  statistics of radio halos.  One of our final data products is a complete
  cosmological cluster mock catalog, at different redshifts, containing
  information of the DM and ICM properties for each object, together with
  its (non-)thermal emission at different frequencies. All the products can
    be found on-line at the MultiDark database (\textit{www.multidark.org}).

    In Section~\ref{sec:2}, we introduce the MultiDark simulation along with our
    selected cluster halo sample. We assign to each of our clusters the
    phenomenologically constructed gas density profile in
    Section~\ref{sec:3}. In Sections~\ref{sec:4} and \ref{sec:5}, we show that
    this approach can successfully reproduce the observed X-ray and SZ cluster
    data. We present the resulting mock cluster catalogs in Section~\ref{sec:6},
    and eventually summarize in Section~\ref{sec:7}.  In this work, the cluster
    mass $M_{\Delta}$ and radius $R_{\Delta}$ are defined with respect to a
    density that is $\Delta=200$ or 500 times the \emph{critical} density of the
    Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.27$,
    $\Omega_{\Lambda}=0.73$ and today's Hubble constant of $H_0 = 100\,
    h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.}

@@

{\bf In Paper~II, we present predictions for the LOFAR radio observatory which
  expects to detect diffuse radio emission in clusters up to redshift $z \approx
  1$ (e.g., \citealp{2012JApA..tmp...34R}).} Thus, we make use of different
simulation snapshots up to $z=1$.  {\bf The extrapolation of our model beyond
  this redshift is rather uncertain as it is based on observations at low(er)
  redshift.} In Table~\ref{tab:z}, we show the total cluster number in our final
cluster sample at different redshifts.

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/lx_m.eps}
\includegraphics[width=0.48\textwidth]{figures/yx_m.eps}
\caption{{\bf X-ray cluster scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result by \protect\cite{2010MNRAS.406.1773M}. 
  The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin. The bottom panels show the relative difference to the observational
    scaling relations. \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}.}} 
\label{fig:X_LM}
\end{figure*}

@@

{\bf So far, we used a well-observed representative sample of X-ray clusters
(REXCESS) that was supplemented with X-ray data on gas mass fractions to construct
our phenomenological gas model.  Here, we will test how our cluster mock data
compare to various X-ray inferred observables, which are obtained from different
cluster samples and from (partially) different X-ray observatories. Those
include the observed $L_{\rmn{bol}} - M_{500}$ relation, the $Y_X - M_{500}$
relation, and the XLF.}

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m.eps}
\includegraphics[width=0.48\textwidth]{figures/sz_m_ACT.eps}
\caption{{\bf Comparison of our MultiDark SZ scaling relations (grey triangles)
    with data. The black crosses represent the $Y_{\rmn{SZ}}$ median values of
    our sample for a given mass bin (indicated by horizontal error bars), and
    the vertical error bars represent the standard deviation within a bin.  The
    bottom panels show the relative difference to the observational scaling
    relations.  \emph{Left.}  We compare our $Y_{\rmn{SZ}}-M$ relation at
    $z=0.18$ with the {\em Planck} data \citep[blue,][]{2013arXiv1303.5080P},
    which have a median redshift of about $0.18$. \emph{Right.} We compare our
    $Y_{\rmn{SZ}}-M$ relation at $z=0.53$ with the ACT sample
    \protect\citep[blue,][]{2013JCAP...07..008H} assuming a self-similar
    redshift evolution.}}
\label{fig:SZ_M}
\end{figure*}

In the left panel of Fig.~\ref{fig:SZ_M}, we compare the $Y_{\rmn{SZ}}-M_{500}$
relation of our model (following \citealp{2012ApJ...758...74B}, equation~3) with
the observed scaling relation by the \cite{2013arXiv1303.5080P}. {\bf We use the
  \emph{Planck} COSMO sample, which is Malmquist-bias corrected and has a median
  redshift of $0.18$, and compare this to our $z=0.18$ relation}. Our model
reproduces the normalization of the scaling relation remarkably well, except for
the high-mass end where our simulations have a weaker constraining power due to
the smaller box size in comparison to the survey volume of {\em Planck}.  {\bf
  However, the slope of the MultiDark $Y_{\rmn{SZ}}-M_{500}$ relation ($1.59\pm0.01$) is
  shallower in comparison to the self-similar slope ($5/3$) as well as the
  observed slope of the \emph{Planck} COSMO sample. We can analytically
  understand the cluster-mass scaling of our model by considering its
  mass-dependent quantities, namely $Y_{\rmn{SZ}}\propto M_{500} f_{\rmn{gas}}
  T \propto M_{500}^{1.595}$. In particular, we can trace back the
  shallower slope to the adopted temperature scaling of our model
  ($T\approx T_{\rmn{ci}}\propto M_{500}^{0.46}$), which includes the cluster core region
  and yields a shallower mass scaling in comparison to the self-similar
  expectation ($T\propto M_{500}^{2/3}$). The temperature scaling is only
  partially countered by the scaling of the gas mass fraction with cluster mass,
  $f_{\rmn{gas}}\propto M_{500}^{0.135}$. To improve upon this, we would have to
  account for a slightly steeper temperature-mass scaling relation (e.g., using a
  centrally excised temperature scaling, see \citealt{2010MNRAS.406.1773M}) in
  the outer cluster parts that are of relevance for the SZ flux, as explained in
  Sect.~\ref{sec:Xscaling}. }  We find a scatter of $\sigma_{yx} \approx 0.11$
which compares well with the \emph{Planck} result of $\sigma_{yx} \approx
0.08$. In Table~\ref{tab:YSZfits}, we report our SZ scaling relations for
different redshifts.

@@

{\bf In the right panel of Fig.~\ref{fig:SZ_M}, we compare our scaling relation
  to the recent Atacama Cosmology Telescope (ACT) sample
  \citep{2013JCAP...07..008H}, which has a median redshift of $z=0.53$.  Again,
  the slope of our modeled scaling relation is slightly shallower in comparison
  to the observed slope. Most strikingly, there appears to be a 30 per cent
  offset of the scaling relations, which could be due to a number of reasons. On
  the data side, these could include possible selection effects of the different
  cluster samples or possible biases associated with density clumping and/or
  non-thermal pressure contributions inherited from the X-ray informed
  post-processing that was employed by the ACT Collaboration to construct their
  $Y_{500}$ estimates. This could also indicate that our model is too simplistic
  to capture the rich cluster physics and potentially may even point to
  deviations from the assumed self-similar redshift evolution (since the ACT
  sample probes on average higher redshift and less massive clusters in
  comparison to the {\em Planck} sample).  However, note that our scaling
  relation at $z=0.53$ ($A=-27.67\pm0.12$, $B=1.60\pm0.01$) is within the
  uncertainties compatible with that of the ACT sample ($A=-28.59\pm2.26$,
  $B=1.65\pm0.15$).  Thus, we conclude that overall there is reasonable
  agreement between our phenomenological model and the SZ data of the
  \emph{Planck} and ACT collaborations.}

@@

{\bf As a final product of this work, we construct cosmologically complete mock
  catalogs of galaxy clusters at different redshifts (see Table~\ref{tab:z}) for
  $M_{200}\geq1\times10^{14}$~$h^{-1}\rmn{M}_{\odot}
  \approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$.  We make these catalogs
  publicly available through the MultiDark database (\textit{www.multidark.org})
  where we will also post possible updates.

  Our catalogs contain all the information regarding the DM properties of each
  cluster as given by the corresponding original MultiDark BDM halo
  catalogs. Additionally, we include the following information:

@@

In Paper~II, we present a model that allows us to compute the possible radio and
gamma-ray emission for each cluster in our mock catalog, as result of the
cosmic-ray (CR) proton-proton interactions with the ICM. We refer to that paper
for all the details. However, we want to point out that our mock catalogs also
include:

@@

We note that our mock catalogs are not appropriate for all purposes because the
density profiles that we adopt in our model represent the average cluster
population of cool-core and non cool-core systems (while we account for scatter
in the gas mass fractions, i.e., the normalization of the profiles). As such,
our mocks are very useful as a baseline for comparisons to new hydrodynamical
simulations and to new observational samples, as well as for non-thermal
modeling of clusters. On the contrary, in the current form they are not
appropriate to perform detailed cosmology studies. Careful modeling of the
response function of the considered instrument, in combination with light-cone
mock catalogs, would be required in this case. Nevertheless, for volume-limited
cluster samples, our mock catalogs will certainly serve as valuable tools. }

@@

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. Note that our model is calibrated
  with X-ray observables within $R_{500}$. More work would be needed to
  extrapolate the model to larger cluster radii that are characterized by an
  increased complexity, which manifests itself in a larger kinetic-to-thermal
  pressure support, an increased level of density clumping, as well as cluster
  asphericity and ellipticity. In our model, we assign a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution. Applying the model to our sample
  of cluster halos, we obtain volume-limited mock catalogs for the thermal (SZ
  and X-ray) cluster emission. Our mock catalogs match the observed X-ray
  luminosity function as well as the $L_{\rmn{X, bol}}-M_{500}$,
  $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$ relations well.

  However, there are some deviations among the different observational scaling
  relations and our model, which may either point to observational sample
  selection effects that are not accounted for or missing complexity of our
  model. This model was specifically constructed to provide a reliable
  description of the gas density, which necessarily implies a great match of the
  resulting mock X-ray observables to data (scaling relations and luminosity
  functions), as well as to the low-redshift and massive cluster sample probed by
  {\em Planck}. However, we note that our model has a slightly shallower slope
  of the $Y_{\rmn{SZ}}-M_{500}$ relation, which implies that the X-ray inferred
  deviations of self-similarity on small scales become weaker on the larger
  scales probed through the SZ effect. Moreover, our model overpredicts the
  normalization of the ACT $Y_{\rmn{SZ}}-M_{500}$ relation (at $z=0.53$) by
  about 30 per cent. While this offset is still consistent within the 1-$\sigma$
  uncertainty intervals of the ACT scaling relation, this could also indicate
  possible selection effects of the different cluster samples or alternatively
  points to a deviation from the assumed self-similar redshift evolution.
